### 2012-02-20 paper.tex
Revised

\noindent \normalsize{High-quality mass spectrometry data of the oscillatory behavior of CO oxidation on silicon dioxide supported Pt-nanoparticles at atmospheric pressure has been performed. The oscillations are self-sustained for several days at constant temperature, pressure and CO/O$_2$ ratio. The frequency of the oscillations is very well defined and increases with time. The oscillation frequency is furthermore strongly temperature dependent with increasing temperature resulting in increasing frequency. A plausible mechanism for the oscillations is proposed based on a oxidation/reduction cycle of the nanoparticles which change the rate of CO oxidation on the particles.}  %on an earlier model on Pd single crystals [Hendriksen \textit{et al., Nature Chemistry}, 2010, \textbf{9}, 730]}
\vspace{0.5cm}
 \end{@twocolumnfalse}
  ]

@@

Here we report on oscillatory behavior of CO oxidation on size-selected platinum nanoparticles supported on a silicon dioxide substrate at atmospheric pressure. Oscillating reactions on platinum at UHV-conditions has been studied extensively and is know to originate from surface reconstructions of the surface during the oscillations\cite{Ertl2008}. This mechanism is, however, not relevant at atmospheric pressures and thus another mechanism is governing the oscillations at high pressures. It has previously been shown\cite{SALES1982} that CO oxidation at atmospheric pressures shows oscillatory behavior on single crystals of Pt, Pd and Ir. Recent studies have proposed a mechanism on palladium \cite{Hendriksen2010} attributed to a cycle of surface oxide formation, oxide roughening and clean metal surface. CO oxidation oscillations has also been observed on Pt thin-films\cite{Lund2000} and in macroscopic flow reactors\cite{Singh2010} but never on size selected platinum nanoparticles at atmospheric pressures for prolonged periods of time.

\section{Experimental Methods}
\subsection{Reactor}
The experiments were performed in silicon-based 20x15 mm $\mu$-reactors\cite{Henriksen2009}. The reactor consists of two inlets, a mixing zone allowing for diffusional mixing of reactants, an outlet, a reactor volume of $240\,$nL and a $5.4\,\mu$m wide capillary used for sniffing gases from the reactor volume. The reactor is sealed by anodic bonding of a pyrex lid to the silicon reactor and is able to operate at a pressure range of 0-2.5\,bar. The reaction gases is supplied to the two inlets by flow controllers capable of controlling the gas flow from 0-10 mL/min. The capillary flow is fed into a quadropole mass-spectrometer for analysis while any surplus of gas from the inlets are passed directly through to the outlet via a pressure controller allowing control of reactor pressure. The design makes sure that all gases exposed to the catalyst under investigation is measured by the mass spectrometer ensuring an extremely high sensitivity of the system.

Before any measurements are performed the reactor is pumped by a turbopump to minimize contaminants in the system. When an evacuated reactor is mounted the base pressure of the mass spectrometer chamber is $\sim5\times10^{-9}\,$mbar and in operation at $1\,$bar in the reactor volume the pressure is $\sim5\times10^{-7}\,$mbar in the mass spectrometer chamber.

The reactor is heated by joule heating of a platinum strip evaporated on the backside of the reactor volume. The introduction of two additional contacts on the backside of the reactor allows for 4 wire measurements of the resistance of the platinum strip using it as a RTD for temperature measurement of the reactor. An external thermocouple acts as room-temperature calibration of the RTD as well as sanity check of the RTD measurement during operation.

@@

\subsection{Platinum nanoparticle deposition}
Platinum nanoparticles were deposited in the reactor using a gas-aggregation source (Mantis Deposition Ltd., Nanogen 50). Platinum clusters were formed by gas-phase condensation of platinum atoms produced by impinging argon ions on a 99.99\% platinum target in a magnetron sputter source. After condensation the ionised fraction (60-80\%) of the clusters were size-selected by a quadropole mass selection filter according to their mass-to-charge ratio. Using this setup platinum nanoparticles with diameters in the range of 2--16\,nm can be produced \cite{Nielsen2010,Nielsen2009}. The size-selected nanoparticles were after size-selection soft-landed in the reactor volume of the microreactor. The coverage was kept at approximately 0.1\% geometric coverage determined by measuring the current on the reactor during deposition. After deposition the reactors were anodicly cold-bonded \cite{Vesborg2010} to a pyrex lid to avoid sintering of the nanoparticles while bonding.

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup directly after cluster deposition and anodic bonding. Initially a light-off CO-oxidation measurement is performed where the temperature is increased at 3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to CO$_2$. At O$_2$:CO ratio of $\sim0.08$ and a geometrical platinum coverage of $\sim0.1\%$ lightoff occurs at approximately 180\,$^\circ$C. The temperature ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is decreased at the same rate as the heating ramp. After the initial lightoff the sample is cooled to room temperature. When at room temperature the temperature is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical lightoff ramp and the subsequent increase in temperature to achieve sustained oscillations is seen. Often the sample will show oscillations either directly upon lightoff or on the falling temperature ramp.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After performing a lightoff ramp where oscillations can been seen the temperature is increased to a constant value where sustained oscillations take place.}
  \label{fgr:initial_treatment}
\end{figure}
The qualitative same behavior has been reproduced on a large number of samples. Nanoparticles with sizes ranging from 3\,nm to 9\,nm, have been tested but no clear size dependence was seen. 3\,nm particles with a geometrical coverage of 0.1\% of the reactor area gave the most stable oscillations. Higher coverages also showed oscillations, but often only on the inital heating and cooling ramps, but only very rarely did they show prolonged oscillations at constant temperature.

@@

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations extracted from the same four day measurement. Data from shortly after the initial treatment, after 2800\,min. of oscillation time and after 5900\,min. is shown. After 5900\,min the oscillations are more irregular which is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also, oscillations in between full conversion oscillations with smaller amplitude and much higher frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 day long experiment. The oscillation period become more irregular with time while the on/off ratio remains constant. Furthermore, small oscillations in between full conversion cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{figure}

The oscillation period becomes more irregular with progressing experiment time although the on/off ratio remains constant during the entire experiment time. Furthermore, the oscillations have the qualitatively identical behavior as shown in Figure~\ref{fgr:full_oscillation}. A short full conversion period is followed by a fast deactivation from where reactivity is slowly recovered. At approximately 50\% conversion lightoff occurs and the sample is in full conversion.

@@

The pressure dependence of the oscillations were also investigated. In the pressure range of 0.1\,bar to 1\,bar no change that could be attributed to the pressure was found. The oscillations do not change frequency nor qualitative behavior at different reactor pressures. Similarly, only a weak dependence on the CO/O$_2$ pressure was observed. At a CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations were observed. Increasing the CO/O$_2$ showed clear oscillations with a significant increase in oscillation amplitude compared to lower CO concentrations. Generally, a tendency towards slower oscillation periods for higher CO/O$_2$ ratio was observed. Richer CO mixtures than 0.175 did not produce any oscillations in our system. This ratio agrees well with literature \cite{Singh2010,Hendriksen2005} where oscillations have been observed in oxygen rich CO/O$_2$ mixtures. The needed low concentration of CO in the inlet gases can be attributed to the much higher sticking coefficient of CO on platinum which poisons the nanoparticle surface during the low reactivity region of the cycle.

Although oscillations have previously been observed on extended platinum thin films \cite{Singh2010} as well as single crystals\cite{Hendriksen2005} at atmospheric pressure no oscillations could be provoked on thin films of comparable geometrical coverage to the nanoparticles in our system at any CO/O$_2$ ratio.

@@

According to the proposed model the bare metal surface is less active than the oxidized surface. It is further suggested due to the general high adsorption energy of CO on metal surfaces that CO stabilizes a metal surface stable at low CO concentrations on high-density stepped surfaces. Contrary, a low concentration of steps will favour an oxidized surface. The oscillating behavior is hence attributed to the shift between an oxidized surface which will increase surface roughness and a reduced surface which will decrease surface roughness. Thus, the reaction will only run a given time in the high-rate state before the catalyst will expose the reduced surface and lower the rate. After some time, the roughness will again be sufficiently low that the catalyst will again oxidize and increase the rate.

In this study Pt nanoparticles were investigated which, compared to a single crystal or a thin film, is a very rough surface. It is to be expected that the particles will initially favor the reduced state. As the reaction runs, the conditions toward the outlet of the reactor will become more oxidizing and at the same time the reaction itself will smoothen the particles further pushing towards an oxidation of the particles. Gradually as the particles toward the outlet oxidizes the turn-over will increase and the general CO concentration will decrease promoting more and more particles to oxidize which will gradually be seen in the mass-spectrometry as an increase in CO$_2$ and a decrease in CO.

@@

>>>>

\noindent \normalsize{High-quality mass spectrometry data of the oscillatory behavior of CO oxidation on SiO$_2$ supported Pt-nanoparticles at atmospheric pressure has been acquired. The oscillations are self-sustained for several days at constant temperature, pressure and CO/O$_2$ ratio. The frequency of the oscillations is very well defined and increases with time. The oscillation frequency is furthermore strongly temperature dependent with increasing temperature resulting in increasing frequency. A plausible mechanism for the oscillations is proposed based on a oxidation/reduction cycle of the nanoparticles which change the rate of CO oxidation on the particles.}  %on an earlier model on Pd single crystals [Hendriksen \textit{et al., Nature Chemistry}, 2010, \textbf{9}, 730]}
\vspace{0.5cm}
 \end{@twocolumnfalse}
  ]

@@

Here we report on oscillatory behavior of CO oxidation on size-selected Pt nanoparticles supported on a SiO$_2$ substrate at atmospheric pressure. Oscillating reactions on Pt at UHV-conditions has been studied extensively and is know to originate from surface reconstructions of the surface during the oscillations\cite{Ertl2008}. This mechanism is, however, not relevant at atmospheric pressures and thus another mechanism is governing the oscillations at high pressures. It has previously been shown\cite{SALES1982} that CO oxidation at atmospheric pressures shows oscillatory behavior on single crystals of Pt, Pd and Ir. Recent studies have proposed a mechanism on palladium \cite{Hendriksen2010} attributed to a cycle of surface oxide formation, oxide roughening and clean metal surface. CO oxidation oscillations has also been observed on Pt thin-films\cite{Lund2000} and in macroscopic flow reactors\cite{Singh2010} but never on size selected Pt nanoparticles at atmospheric pressures for prolonged periods of time.

\section{Experimental Methods}
\subsection{Reactor}
The experiments were performed in Si-based 20x15 mm $\mu$-reactors\cite{Henriksen2009}. The reactor consists of two inlets, a mixing zone allowing for diffusional mixing of reactants, an outlet, a reactor volume of $240\,$nL and a $5.4\,\mu$m wide capillary used for sniffing gases from the reactor volume. The reactor is sealed by anodic bonding of a pyrex lid to the Si reactor and is able to operate at a pressure range of 0-2.5\,bar. The reaction gases is supplied to the two inlets by flow controllers capable of controlling the gas flow from 0-10 mL/min. The capillary flow is fed into a quadropole mass-spectrometer (QMA) for analysis while any surplus of gas from the inlets are passed directly through to the outlet via a pressure controller allowing control of reactor pressure. The design makes sure that all gases exposed to the catalyst under investigation is measured by the QMA ensuring an extremely high sensitivity of the system.

Before any measurements are performed the reactor is pumped by a turbopump to minimize contaminants in the system. When an evacuated reactor is mounted the base pressure of the mass spectrometer chamber is $\sim5\times10^{-9}\,$mbar and in operation at $1\,$bar in the reactor volume the pressure is $\sim5\times10^{-7}\,$mbar in the QMA chamber.

The reactor is heated by joule heating of a Pt strip evaporated on the backside of the reactor volume. The introduction of two additional contacts on the backside of the reactor allows for 4 wire measurements of the resistance of the Pt strip using it as a RTD for temperature measurement of the reactor. An external thermocouple acts as room-temperature calibration of the RTD as well as sanity check of the RTD measurement during operation.

@@

\subsection{Pt nanoparticle deposition}
Pt nanoparticles were deposited in the reactor using a gas-aggregation source (Mantis Deposition Ltd., Nanogen 50). Pt clusters were formed by gas-phase condensation of Pt atoms produced by impinging argon ions on a 99.99\% Pt target in a magnetron sputter source. After condensation the ionised fraction (60-80\%) of the clusters were size-selected by a quadropole mass selection filter according to their mass-to-charge ratio. Using this setup Pt nanoparticles with diameters in the range of 2--16\,nm can be produced \cite{Nielsen2010,Nielsen2009}. The size-selected nanoparticles were after size-selection soft-landed in the reactor volume of the microreactor. The coverage was kept at approximately 0.1\% geometric coverage determined by measuring the current on the reactor during deposition. After deposition the reactors were anodicly cold-bonded \cite{Vesborg2010} to a pyrex lid to avoid sintering of the nanoparticles while bonding.

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup directly after cluster deposition and anodic bonding. Initially a light-off CO-oxidation measurement is performed where the temperature is increased at 3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to CO$_2$. At O$_2$:CO ratio of $\sim0.08$ and a geometrical Pt coverage of $\sim0.1\%$ light-off occurs at approximately 180\,$^\circ$C. The temperature ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is decreased at the same rate as the heating ramp. After the initial light-off the sample is cooled to room temperature. When at room temperature the temperature is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical light-off ramp and the subsequent increase in temperature to achieve sustained oscillations is seen. Often the sample will show oscillations either directly upon light-off or on the falling temperature ramp.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After performing a light-off ramp where oscillations can been seen the temperature is increased to a constant value where sustained oscillations take place.}
  \label{fgr:initial_treatment}
\end{figure}
The qualitative same behavior has been reproduced on a large number of samples. Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm particles with a geometrical coverage of 0.1\% of the reactor area gave the most stable oscillations. Higher coverages also showed oscillations but often only on the inital heating and cooling ramps. Only on rare occasions sustained oscillations were measured at constant temperature.

@@

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations extracted from the same four day measurement. Data from shortly after the initial treatment, after 2800\,min. of oscillation time and after 5900\,min. is shown. After 5900\,min. the oscillations are more irregular which is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also, oscillations in between full conversion oscillations with smaller amplitude and much higher frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 day long experiment. The oscillation period become more irregular with time while the on/off ratio remains constant. Furthermore, small oscillations in between full conversion cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{figure}

The oscillation period becomes more irregular with progressing experiment time although the on/off ratio remains constant during the entire experiment time. Furthermore, the oscillations have the qualitatively identical behavior as shown in Figure~\ref{fgr:full_oscillation}. A short full conversion period is followed by a fast deactivation from where reactivity is slowly recovered. At approximately 50\% conversion light-off occurs and the sample is in full conversion.

@@

The pressure dependence of the oscillations were also investigated. In the pressure range of 0.1\,bar to 1\,bar no change of the oscillation frequency or qualitative behavior that could be attributed to the pressure was found. Similarly, only a weak dependence on the CO/O$_2$ ratio was observed. At a CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations were observed. Increasing the CO/O$_2$ showed clear oscillations with a significant increase in oscillation amplitude compared to lower CO concentrations. Generally, a tendency towards slower oscillation periods for higher CO/O$_2$ ratio was observed. Richer CO mixtures than 0.175 did not produce any oscillations in our system.

Although oscillations have previously been observed on extended Pt thin films \cite{Singh2010} as well as single crystals\cite{Hendriksen2005} at atmospheric pressure no oscillations could be provoked on thin films of comparable geometrical coverage to the nanoparticles in our system at any CO/O$_2$ ratio.

@@

According to the proposed model the bare metal surface is less active than the oxidized surface. It is further suggested due to the general high adsorption energy of CO on metal surfaces that CO stabilizes a metal surface stable at low CO concentrations on high-density stepped surfaces. Contrary, a low concentration of steps will favour an oxidized surface and thus high activity. The oscillating behavior is hence attributed to the shift between an oxidized surface which will increase surface roughness and a reduced surface which will decrease surface roughness and hence the rate. Thus, the reaction will only run a given time in the high-rate state before the catalyst will expose the reduced surface and lower the rate. After some time, the roughness will again be sufficiently low that the catalyst will again oxidize and increase the rate.

In this study Pt nanoparticles were investigated which, compared to a single crystal or a thin film, is a very rough surface. It is to be expected that the particles will initially favor the reduced state. As the reaction runs, the conditions toward the outlet of the reactor will become more oxidizing and at the same time the reaction itself will smoothen the particles further pushing towards an oxidation of the particles. Gradually as the particles toward the outlet oxidizes the turn-over will increase and the general CO concentration will decrease promoting more and more particles to oxidize which will gradually be seen in the QMA as an increase in CO$_2$ and a decrease in CO.

@@

Oscillations were only seen in CO/O$_2$ ratio below 0.175. This ratio agrees well with literature \cite{Singh2010,Hendriksen2005} where oscillations have been observed in oxygen rich CO/O$_2$ mixtures. The needed low concentration of CO in the inlet gases can be attributed to the much higher sticking coefficient of CO on Pt which poisons the nanoparticle surface during the low reactivity region of the cycle.
